---
title: "L1FR-L2ES-Effect"
author: "CMVA"
date: "2024-11-07"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(stringr)
library(dplyr)
library(ggplot2)
library(readxl)
library(purrr)
```

# Initial Read-in

The data was downloaded from the corpus originally as separate excel files that were all A1 level Spanish learners, but each file was for a different conjugation of the verb *ir* 'to go'.

There are two datasets: L1 French speakers and L1 English speakers. These were chosen in order to analyze whether having a language with the same periphrastic future construction does actually affect language transfer.

The individual Excel files will be read in and then combined into one dataset (per language) for analysis.

```{r}
getwd()
```

```{r read-in}
read_excel("Excels/A1IR.xlsx")
read_excel("Excels/A1VOY.xlsx")
read_excel("Excels/A1VAS.xlsx")
read_excel("Excels/A1VA.xlsx")
read_excel("Excels/A1VAMOS.xlsx")
read_excel("Excels/IRENG.xlsx")
read_excel("Excels/VOYENG.xlsx")
read_excel("Excels/VASENG.xlsx")
read_excel("Excels/VAENG.xlsx")
read_excel("Excels/VAMOSENG.xlsx")
read_excel("Excels/VANENG.xlsx")
```
```{r combine}
FRENG <- list.files(path = "Excels", pattern = "*.xlsx*", all.files = TRUE, full.names = TRUE, recursive = FALSE)
A1 <- FRENG %>%
    set_names() %>%
    map_dfr(read_excel, .id = "Source")
A1

write.csv(A1, "A1.csv", row.names = FALSE)
```

I am starting by translating all of the columns into English for ease of understanding and access for learners who are not proficient in Spanish.
```{r}
A1IR <- rename(A1, Student = Estudiante,
       Task = Tarea,
       TypeText = `Tipología textual`,
       Theme = Tema,
       Age = Edad,
       Gender = Sexo,
       Country = País,
       Studies = Estudios,
       AgeStart = `Edad de inicio en el estudio del español`,
       Months = `Número de meses estudiando español`,
       Contacts = `Contactos personales en países de habla española`,
       ContextL = `Contexto a la izquierda`,
       GramElem1 = `Elemento gramatical 1`,
       Tag1 = `Etiqueta 1`,
       Lemma1 = `Lema 1`,
       OrthoWord1 = `Palabra ortográfica 1`,
       ExInfo1 = `Información adicional 1`,
       ContextR = `Contexto a la derecha`)
```
# Tidying

The first thing I want to do with my datasets is see if there are any duplicate productions from the same participant within the different conjugations of "voy".
```{r duplicate FR}
A1IR
unique(A1IR$Student)
count(A1IR,Student)

A1IRNEW <- A1IR %>%
  group_by(Student)
A1IRNEW

A1IRNEW %>% 
  filter(L1 == "Francés")
A1IRNEW %>% 
  filter(L1 == "Inglés")
```
Based on the information above, we see that there are a total of 48 students with French as their L1, with each participant having a different number of total productions.As for English, there are 35 total students with English as their L1. In total, we have 83 total students in this dataset.

The second thing I am going to do is remove any columns I deem unnecessary for my analyses. These include the 17th and 18th columns, as they have no data.
```{r column removal}
A1 <- subset(A1IRNEW, select = -c(OrthoWord1, ExInfo1))
A1
```

# Basic Statistics

Now, I am going to do some basic statistical inquiries, just to see the range of ages, as well as what country they are from, how old they were when they started (to calculate how long they have been learning Spanish), etc.

```{r contacts FR}
count(A1, Contacts)
unique(A1$Contacts)
A1 %>% 
  filter(Contacts == "No")
A1 %>% 
  filter(Contacts == "Amigos")
A1 %>% 
  filter(Contacts == "Amigos&Familiares")
A1 %>% 
  filter(Contacts == "Familiares")
A1 %>% 
  filter(Contacts == "Otros")
```
The contacts column is meant for the students to relay whether they have any contacts who are Spanish speakers. There were a total of five distinct answers given from the 48 L1 French students:

1. No contacts
2. "Amigos" -> Friends
3. "Amigos&Familiares" -> Friends and Family
4. "Familiares" -> Family
5. "Otros" -> Others

The data was filtered five times for each category of answers to how many tokens were accredited to each category of answers. The results show that the majority of students with L1 French had no contacts who are Spanish speakers (total of 25 students, 52%). The category with the smallest students is the "Friends" category, with a total of 10 students stating they have family members who are Spanish speakers, adding up to approximately 10% of the students in the data.

As for students with an L1 of English, there were a total of four distinct answers given from the 35 students:

1. No contacts
2. "Amigos" -> Friends
3. "Amigos&Familiares" -> Friends and Family
4. "Otros" -> Others

The results show that the majority of students with L1 English had no contacts who are Spanish speakers (total of 21 students, 60%). The category with the smallest students is the "Friends and Family" category, with a total of 2 students stating they have friends and family members who are Spanish speakers, adding up to approximately 5.7% of the students in the data.

```{r themes}
unique(A1$Theme)
```
The CAES corpus is comprised of all written texts from individuals who are L2 learners of Spanish. For these particular datasets, there were three distinct topics that the students wrote about: Job change, Family, and Note of late arrival.

```{r type text}
unique(A1$TypeText)
```
There are two types of written data by the students in each dataset:

1. "Correo electronico" -> E-mail
2. "Nota/aviso" -> Note/Warning

```{r country}
count(A1, Country)
unique(A1$Country)
```

```{r studies}
count(A1, Studies)
```

```{r age FR}
mean(A1$Age)

A1AGE <- A1 %>% 
  group_by(Age) %>% 
  summarise(count = n())
  print(A1AGE)
```
As can be seen above, it appears that ages in the L1 French dataset range from 15 to 71 years old. Not only that, but the age group with the most amount of students is the 61 year-olds who have 13 total individuals. The age group with the least amount of students is the 15, 18, 19, 24, and 44 year-olds who only had one individual per age. (*need to edit this for ENG*)

# Picking Out the Prescriptive Output

```{r A1 a}
CA1 <- A1 %>%
    mutate(First2 = sub("^([[:alpha:]]+ [[:alpha:]]+).*", "\\1", ContextR))


# If needed, save the modified data frame to a new object or overwrite the existing one
A1 <- CA1
A1
```

# Statistical Analyses

```{r model 1}

A1 <- A1 %>% 
  mutate(Correct = ifelse(grepl("^a.*r$", First2), 1, 0))

print(A1)

summary_df <- A1 %>%
    group_by(Age, Gender, Country) %>%
    summarise(CorrectAnswers = sum(Correct))

print(summary_df)

model1 <- glm(Correct ~ Age + Gender + Country, data = A1, family = binomial)

summary(model1)
```

```{r model 2}
model2 <- glm(Correct ~ L1 + AgeStart + Months + Contacts, data = A1, family = binomial)

summary(model2)
```
```{r model 3}
model3 <- glm(Correct ~ L1 + Country, data = A1, family = binomial)

summary(model3)
```
# Visualizations

```{r}
```

```{r}
count(A1, Correct)
unique(A1$Correct)
```

```{r plot 1}
ggplot(A1, aes(x = Correct, y = Country)) +
  geom_point(color = "red", size = 1) +
  labs(title = "Scatter Plot of Country vs. Correctness", x = "Country", y = "Correctness") + 
  theme_minimal()
```


```{r session info}
library(sessioninfo)
session_info()
```

