---
title: "L1FR-L2ES-Effect"
author: "CMVA"
date: "2024-11-07"
output: github_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(stringr)
library(dplyr)
library(ggplot2)
library(readxl)
library(purrr)
```

# Initial Read-in

The data was downloaded from the corpus originally as separate excel files that were all A1 level Spanish learners, but each file was for a different conjugation of the verb *ir* 'to go'.

There are two datasets: L1 French speakers and L1 English speakers. These were chosen in order to analyze whether having a language with the same periphrastic future construction does actually affect language transfer.

The individual Excel files will be read in and then combined into one dataset (per language) for analysis.

```{r}
getwd()
```

```{r read-in}
read_excel("Excels/A1IR.xlsx")
read_excel("Excels/A1VOY.xlsx")
read_excel("Excels/A1VAS.xlsx")
read_excel("Excels/A1VA.xlsx")
read_excel("Excels/A1VAMOS.xlsx")
read_excel("Excels/IRENG.xlsx")
read_excel("Excels/VOYENG.xlsx")
read_excel("Excels/VASENG.xlsx")
read_excel("Excels/VAENG.xlsx")
read_excel("Excels/VAMOSENG.xlsx")
read_excel("Excels/VANENG.xlsx")
```
```{r combine}
FRENG <- list.files(path = "Excels", pattern = "*.xlsx*", all.files = TRUE, full.names = TRUE, recursive = FALSE)
A1 <- FRENG %>%
    set_names() %>%
    map_dfr(read_excel, .id = "Source")
A1

write.csv(A1, "A1.csv", row.names = FALSE)
```

I am starting by translating all of the columns into English for ease of understanding and access for learners who are not proficient in Spanish.
```{r}
A1IR <- rename(A1, Student = Estudiante,
       Task = Tarea,
       TypeText = `Tipología textual`,
       Theme = Tema,
       Age = Edad,
       Gender = Sexo,
       Country = País,
       Studies = Estudios,
       AgeStart = `Edad de inicio en el estudio del español`,
       Months = `Número de meses estudiando español`,
       Contacts = `Contactos personales en países de habla española`,
       ContextL = `Contexto a la izquierda`,
       GramElem1 = `Elemento gramatical 1`,
       Tag1 = `Etiqueta 1`,
       Lemma1 = `Lema 1`,
       OrthoWord1 = `Palabra ortográfica 1`,
       ExInfo1 = `Información adicional 1`,
       ContextR = `Contexto a la derecha`)
```
# Tidying

The first thing I want to do with my datasets is see if there are any duplicate productions from the same participant within the different conjugations of "voy".
```{r duplicate FR}
A1IR
unique(A1IR$Student)
count(A1IR,Student)

A1IRNEW <- A1IR %>%
  group_by(Student)
A1IRNEW

A1IRNEW %>% 
  filter(L1 == "Francés")
A1IRNEW %>% 
  filter(L1 == "Inglés")
```
Based on the information above, we see that there are a total of 48 students with French as their L1, with each participant having a different number of total productions.As for English, there are 35 total students with English as their L1. In total, we have 83 total students in this dataset.

The second thing I am going to do is remove any columns I deem unnecessary for my analyses. These include the 17th and 18th columns, as they have no data.
```{r column removal}
A1 <- subset(A1IRNEW, select = -c(OrthoWord1, ExInfo1))
A1
```

# Basic Statistics

Now, I am going to do some basic statistical inquiries, just to see the range of ages, as well as what country they are from, how old they were when they started (to calculate how long they have been learning Spanish), etc.

```{r contacts FR}
count(A1, Contacts)
unique(A1$Contacts)
A1 %>% 
  filter(Contacts == "No")
A1 %>% 
  filter(Contacts == "Amigos")
A1 %>% 
  filter(Contacts == "Amigos&Familiares")
A1 %>% 
  filter(Contacts == "Familiares")
A1 %>% 
  filter(Contacts == "Otros")
```
The contacts column is meant for the students to relay whether they have any contacts who are Spanish speakers. There were a total of five distinct answers given from the 48 L1 French students:

1. No contacts
2. "Amigos" -> Friends
3. "Amigos&Familiares" -> Friends and Family
4. "Familiares" -> Family
5. "Otros" -> Others

The data was filtered five times for each category of answers to how many tokens were accredited to each category of answers. The results show that the majority of students with L1 French had no contacts who are Spanish speakers (total of 25 students, 52%). The category with the smallest students is the "Friends" category, with a total of 10 students stating they have family members who are Spanish speakers, adding up to approximately 10% of the students in the data.

As for students with an L1 of English, there were a total of four distinct answers given from the 35 students:

1. No contacts
2. "Amigos" -> Friends
3. "Amigos&Familiares" -> Friends and Family
4. "Otros" -> Others

The results show that the majority of students with L1 English had no contacts who are Spanish speakers (total of 21 students, 60%). The category with the smallest students is the "Friends and Family" category, with a total of 2 students stating they have friends and family members who are Spanish speakers, adding up to approximately 5.7% of the students in the data.

```{r themes}
unique(A1$Theme)
```
The CAES corpus is comprised of all written texts from individuals who are L2 learners of Spanish. For these particular datasets, there were three distinct topics that the students wrote about: Job change, Family, and Note of late arrival.

```{r type text}
unique(A1$TypeText)
```
There are two types of written data by the students in each dataset:

1. "Correo electronico" -> E-mail
2. "Nota/aviso" -> Note/Warning

```{r country}
count(A1, Country)
unique(A1$Country)
```

```{r studies}
count(A1, Studies)
```

```{r age}
mean(A1$Age)

A1AGE <- A1 %>% 
  group_by(Age) %>% 
  summarise(count = n())
  print(A1AGE)
```
As can be seen above, it appears that ages in the dataset range from 15 to 71 years old. Not only that, but the age group with the most amount of students is the 61 year-olds who have 13 total individuals. The age group with the least amount of students is the 15, 18, 19, 24, and 44 year-olds who only had one individual per age.

# Picking Out the Prescriptive Output

```{r A1 a}
CA1 <- A1 %>%
    mutate(First2 = sub("^([[:alpha:]]+ [[:alpha:]]+).*", "\\1", ContextR))


# If needed, save the modified data frame to a new object or overwrite the existing one
A1 <- CA1
A1
```

# Statistical Analyses

```{r model 1}
A1 <- A1 %>% 
  mutate(Correct = ifelse(grepl("^a.*r$", First2), 1, 0))

print(A1)

summary_df <- A1 %>%
    group_by(Age, Gender, Country, L1) %>%
    summarise(CorrectAnswers = sum(Correct))

print(summary_df)

model1 <- glm(Correct ~ Age + AgeStart + Gender + Country + Months + Studies + Contacts + L1, data = A1, family = binomial)

summary(model1)
```

## Analysis - Model 1

Significant factors: Age, AgeStart (most), Being from Canada, Starting in Primary School, Starting in Uni (more than Primary)

Almost Significant: Being from South Korea (crazy!), having family members who speak Spanish, and having English as an L1.

# Visualizations

```{r plot 1}
ggplot(model1, aes(AgeStart)) +
  geom_bar(color = "red", fill = "black", size = .5) +
  labs(title = "Bar Plot of Age Start vs. Correctness", x = "Age", y = "Correctness") + 
  theme_minimal()
```
Plot 1 shows that the largest number of correctness came from those who started learning Spanish from birth.

```{r plot 2}
ggplot(model1, aes(Age)) +
  geom_bar(color = "red", fill = "black", size = .5) +
  labs(title = "Bar Plot of Age vs. Correctness", x = "Age", y = "Correctness") + 
  theme_minimal()
```
Plot 2 shows that those who were around 50 years old had the most correct answers.

```{r plot 3}
ggplot(model1, aes(Country)) +
  geom_bar(width = 0.5, color = "red", fill = "black", size = .5) +
  labs(title = "Bar Plot of Country vs. Correctness", x = "Country", y = "Correctness") +
  theme(axis.text.x = element_text(angle = 90)) 
```
Plot 3 shows that those who are from France had the most correct productions of the periphrastic future in French. Although interesting, it was Canada that had a significant effect on the correctness according to the logistic regression.

```{r plot 4}
ggplot(model1, aes(Studies)) +
  geom_bar(width = 0.5, color = "red", fill = "black", size = .5) +
  labs(title = "Bar Plot of Where Students Started Studying Spanish vs. Correctness", x = "When Started", y = "Correctness") +
  theme_minimal()
```
Plot 4 shows that those who started learning Spanish at their university had the most correct answers. The regression analysis shows that both those who started in primary school and at their university had the most significant effect on their correctness, with university learners having slightly more significance (about 0.02 more significance).
```{r plot 5}
ggplot(model1, aes(L1)) +
  geom_bar(width = 0.5, color = "red", fill = "black", size = .5) +
  labs(title = "Bar Plot of L1 vs. Correctness", x = "L1", y = "Correctness") +
  theme_minimal()
```
Plot 5 shows that those whose L1 is French had more correct answers. This actually answers my main research question, which was whether having French as an L1 allows one to have more correct productions than English speakers. According to the regression, having English as an L1 was *almost* significant. The negative estimate for L1 English means that having an L1 of English reduced the production of correct periphrastic future sentences.

```{r plot 6}
ggplot(model1, aes(Contacts)) +
  geom_bar(width = 0.5, color = "red", fill = "black", size = .5) +
  labs(title = "Bar Plot of Contacts Who Speak Spanish vs. Correctness", x = "Contacts", y = "Correctness") +
  theme_minimal()
```
Plot 6 shows that participants with no contact who speaks Spanish had the most correct periphrastic future productions. According to the regression analysis, those with family members who speak Spanish almost had a significant effect on correctness. Its negative estimate means that having family members who speak Spanish reduces the correct productions, which is interesting, considering that should actually be more helpful. But with consideration, they may have a different, less prescriptive production of Spanish which may in turn confuse the participant.

# Conclusion

Based on the information provided above, both research questions were answered. First of all, having French as an L1 *does* in fact help participants produce the periphrastic future better than those who have English as an L1. As for the second question, the factors that affect periphrastic future production in L2 Spanish are the age when they started learning, their current age, whether or not they have any contacts who speak Spanish, what country they are from, what their L1 is, and what school they were in when they learned/started learning Spanish. The age when they started learning proved to be the most significant factor based on the logistic regression model.

# Revisions After Professor Input - 12/11/24, 8:41p

I need to revise the dataset so that we have a Correct and Incorrect column. The reason for this is that knowing what is incorrect allows us to see what is prescriptive. Once I do that, I will double-check all columns to make sure that the Correct indications do not include anything that has "ir" in the word, such as "Irlandes" which means "Irish" and not "to go". 

Once I have two columns, one for Correct and one for Incorrect, then I will filter out anyone who started learning Spanish at birth. This is because they would be considered L1/Native speakers of Spanish if they started at birth, not L2/Second language learners. I am choosing to filter it out for anyone who started learning at age 5 or later, because Musliadi (2019) states that by five years old, a child understands the semantics of language (p. 167). Anyone trying to produce language or make conscious grammatical decisions, such as using the periphrastic future, should be able to process and produce it correctly by the age of five. Their L1 should be largely acquired by then, so the introduction of another language would not be native/bilingual, but an L2. 

```{r}
(A1)
```


```{r incorrect}
A1 <- A1 %>%
  mutate(Incorrect = if_else(Correct == 0, 1, 0))

print(A1)

A1 <- A1 %>% 
  filter(AgeStart >= 5)

print(A1)

table(A1$Correct)
table(A1$Incorrect)
  

library(lme4)

# Ensure categorical variables are properly converted to factors
A1 <- A1 %>%
  mutate(
    L1 = factor(L1),
    Gender = factor(Gender),
    Country = factor(Country),
    Studies = factor(Studies),
    Student = factor(Student)
  )

model2 <- glm(Correct ~ Age + AgeStart + Gender + Country + Months + Studies + Contacts + L1, data = A1, family = binomial)

summary(model2)
```

```{r sk filter}
A1KOR <- A1 %>% 
  filter(Country == "Corea del Sur")

print(A1KOR)
```

```{r model 2 gender}
ggplot(model2, aes(Gender)) +
  geom_bar(color = "red", fill = "black", size = .5) +
  labs(title = "Bar Plot of Gender vs. Correctness", x = "Gender", y = "Correctness") + 
  theme_minimal()
```

The above bar graph shows that participants who identified as women had higher rates of correct productions of the periphrastic future compared to their male counterparts. According to the binomial logistic regression, women specifically had a positive estimate that significantly affected periphrastic future production. Although a small estimate, because it is positive, it demonstrates that being a woman results in more correct productions of periphrastic future in L2 Spanish.

```{r model2 country}
ggplot(model2, aes(Country)) +
  geom_bar(color = "red", fill = "black", size = .5) +
  labs(title = "Bar Plot of Country of Origin vs. Correctness", x = "Country", y = "Correctness") + 
  theme_minimal()+
  theme(axis.text.x = element_text(angle = 90)) 
```

```{r model2 months}
ggplot(model2, aes(Months)) +
  geom_bar(color = "red", fill = "black", size = .5) +
  labs(title = "Bar Plot of Months Spent Studying vs. Correctness", x = "Months", y = "Correctness") + 
  theme_minimal()
```

```{r model2 studies}
ggplot(model2, aes(Studies)) +
  geom_bar(color = "red", fill = "black", size = .5) +
  labs(title = "Bar Plot of School Where Spanish Was Studied vs. Correctness", x = "Studies", y = "Correctness") + 
  theme_minimal()
```

```{r model2 contacts}
ggplot(model2, aes(Contacts)) +
  geom_bar(color = "red", fill = "black", size = .5) +
  labs(title = "Bar Plot of Contacts Who Speak Spanish vs. Correctness", x = "Contacts", y = "Correctness") + 
  theme_minimal()
```


```{r session info}
library(sessioninfo)
session_info()
```

